---
title: "combineTools_v2_neg"
author: "Krista Longnecker"
date: "October 12, 2017"
output: html_document
---

```{r setStage, eval = TRUE, results = 'hide'}

library(ggplot2)
library(XLConnect)
library(reshape2)

source("FindAllWithinError.r")

```


This assumes that all the time consuming calculations have already been run and are the 'right' answer. 
--> run Lumos_MDexp1b_pos_mspurity to get the aligned XCMS data
--> run Lumos_MDexp1b_pos_miner to get the MS2 fragments for each sample (as RDdata files)

One down side to this setup is that if I click 'knit' above, I can only get an HTML file with the same name as the Rmd file...
```{r getData, eval = TRUE}

ionMode <- 'neg'

if (ionMode=='neg') {
  load("MDv1b_mspurity_neg_2017.10.19_v1.Rimage")
  NameOfFile <- 'Lumos_MDv1b_combined_neg.2017.10.19.Rimage'
  saveFigures = 'combineTools_neg_figures/'
  
  mtab_names = cbind("biotin","citric acid",
                    "l-malic acid","orotic acid",
                    "succinic acid","taurocholic acid","thymidine")
                    
  
  mtab_targets = c(243.080885877909,
  191.019726127909,
  133.014246827909,
  155.009830147909,
  117.019332207909,
  514.284396227909,
  241.082995067909)
  
} else if (ionMode == 'pos') {
  load("MDv1b_mspurity_pos_2017.10.14_v1.Rimage")
  
  #set new file name for combined file:
  NameOfFile <- 'Lumos_MDv1b_combined_pos.2017.10.20.Rimage'
  saveFigures = 'combineTools_pos_figures/'
  
  mtab_names = cbind("biotin","caffeine","l-arginine","l-glutamine",
                      "l-methionine","l-proline","l-serine","l-threonine",
                      "n-acetyl-d-glucosamine")

  #hack from list of metabolites on aston (working copy in Lumos folder)
  mtab_targets = c(245.0954388,
                 195.087652,
                 175.1189521,
                 147.0764186,
                 150.058325,
                 116.070605,
                 106.0498695,
                 120.0655196,
                 222.0972136)

}
  
  
  save.image(NameOfFile)

```


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  fig.path = saveFigures)
```


```{r getSampleInfo, eval = TRUE}

#use the sequence file to get the sample information for all the samples
wDir <-  "C:/Users/krista/Documents/Current projects/Kujawinski_Lumos/RawData/2017_0719 Method Development 1_batch 2/sequenceFiles_fromMethods/"
all <- readWorksheetFromFile(paste0(wDir,"mtab_MethodDev_071917_sequence_batch2.KL.xlsx"),sheet = 1)
putDataHere <-data.frame()


for(i in 1:length(all$File.Name)) {
  h <- all$File.Name[i]
  #paste0 will strip out the spaces; make the full filename here    
  h <-paste0(pre,h,ext)
  
  #this will pick up the right row in mzdatafiles
  m <- match(basename(h),basename(xs@filepaths))
  if (!is.na(m)) {
    tData <- all[i,]
    if (nrow(putDataHere)==0) {
      putDataHere <- tData
      } else {
    putDataHere[m,] <-tData
    }
    rm(tData)
  }   
  rm(h,m)
}   

rm(all) #keep this clean...

```


```{r setupDataStorage, eval = TRUE}

#df <- data.frame(number = seq.int(length(mzdatafiles)))
calcVars <- c('nMZ','nMS2','avg_scans_perPrecursor','avg_peaks_per_MS2')

df <- data.frame(matrix(ncol=length(calcVars),nrow=length(mzdatafiles)))
rownames(df) <- basename(mzdatafiles)
colnames(df) <- c(calcVars)

#match up the rows of the data frame with putDataHere to get ancillary information 
#multiple step process in R it seems...(though for this example idx1==idx2)
#first find the intersect
z <- intersect(rownames(df),paste0(putDataHere$File.Name,'.mzML'))
#want the index into rownames(df) 
idx1 <- match(z,rownames(df))
#then find the index into putDataHere
idx2 <- match(z,paste0(putDataHere$File.Name,'.mzML'))


```

##msPurity: general parameters

```{r numberMZ, eval = TRUE, fig.height = 8,fig.width = 10}

#probably a better way to do this....

for (i in 1:length(mzdatafiles)) {
  
  one <- as.data.frame(xs@peaks)
  temp <- one[one$sample == i,]
  
  df$nMZ[i] = dim(temp)[1]
  rm(one,temp)
}

#plot.new()
par(oma = c(5,5,5,5))
x = putDataHere$File.Name[idx2]
y = df$nMZ[idx1]

#don't really need a new data.frame
d <- data.frame(x=x, y=y, collision = putDataHere$cType[idx1],
                energy = putDataHere$cEnergy[idx1], detection = putDataHere$dType[idx1])

# p <- ggplot(d, aes(x=x, y=y, colour=interaction(treatment,replicate))) +
#   geom_point(aes(shape = treatment), size=10)
p <- ggplot(d, aes(x=x, y=y, colour=interaction(collision,energy))) +
  geom_point(aes(shape = detection), size=8)
p <- p + scale_color_brewer(palette = 'Set3')

p <- p + labs(y = "# of MZ values")
#p <- p + scale_x_continuous(breaks=c(1:dim(df)[1]),labels = putDataHere$FileName[idx2])
p <- p + theme_bw()
p <- p +  theme(plot.background = element_blank(),
   panel.grid.major = element_blank())
# p <- p +  theme(plot.background = element_blank(),
#    panel.grid.major = element_blank(),
#    panel.grid.minor = element_blank(),
#    panel.border = element_blank())
p <- p + theme(axis.text.x=element_text(angle = -45, hjust = 0))
p
   

```


```{r numberMS2, eval = TRUE, fig.height = 8,fig.width = 10}

#count # of MS2 for each datafile...find the number of rows for each datafile and get length

for (i in 1:length(mzdatafiles)) {
  w <- which(pa@puritydf$fileid==i)
  df$nMS2[i] = length(w)
  rm(w)

}
  
#I think this is best:
x = putDataHere$File.Name[idx2]
y = df$nMS2[idx1]

d <- data.frame(x=x, y=y, collision = putDataHere$cType[idx1],
                energy = putDataHere$cEnergy[idx1], detection = putDataHere$dType[idx1])

p <- ggplot(d, aes(x=x, y=y, colour=interaction(collision,energy))) +
  geom_point(aes(shape = detection), size=8)
p <- p + scale_color_brewer(palette = 'Set3')

p <- p + labs(y = "# of MS2 scans")
p <- p + theme_bw()
p <- p +  theme(plot.background = element_blank(),
   panel.grid.major = element_blank())
p <- p + theme(axis.text.x=element_text(angle = -45, hjust = 0))
p
   
```



```{r avg_per_percursor, eval = TRUE, fig.height = 8,fig.width = 10}
#Average #of MS2 scans per MS1 scan
#get average value (# peaks) for each file separately and put into df

for (ii in 1:length(mzdatafiles)) {
  ww = which(pa@puritydf$fileid==ii)
  tData <- pa@puritydf[ww,] #want ww rows, all columns
  
  npre = as.data.frame(unique(tData$precursorScanNum))

  for (i in 1:dim(npre)[1]) {
    w = which(tData$precursorNearest==npre[i,1])
    npre[i,2] <- length(w)
    rm(w)
  }
  
  #now get the average for this file
  df$avg_scans_perPrecursor[ii] <- mean(npre$V2)
  rm(ww, tData,npre)
    
}

#plot up the results
#plot.new()
x = putDataHere$File.Name[idx2]
y = df$avg_scans_perPrecursor[idx1]

d <- data.frame(x=x, y=y, collision = putDataHere$cType[idx1],
                energy = putDataHere$cEnergy[idx1], detection = putDataHere$dType[idx1])

p <- ggplot(d, aes(x=x, y=y, colour=interaction(collision,energy))) +
  geom_point(aes(shape = detection), size=8)
p <- p + scale_color_brewer(palette = 'Set3')

p <- p + labs(y = "average # of MS2 scans per MS1")
p <- p + theme_bw()
p <- p +  theme(plot.background = element_blank(),
   panel.grid.major = element_blank())
p <- p + theme(axis.text.x=element_text(angle = -45, hjust = 0))
p


```

##msPurity: only with the standards


```{r iPurity_score, eval = TRUE, fig.height = 4,fig.width = 5}

#export the mz values: this does into puritydf (missed that yesterday)
oneList <- pa@puritydf[,'precursorMZ'] #data.frame syntax
maxErr <- 2
  
mean_iPurity <- data.frame(matrix(ncol=length(mtab_targets),nrow=length(mzdatafiles)))
rownames(mean_iPurity) <- basename(mzdatafiles)
colnames(mean_iPurity) <- c(mtab_names)

#i <- 1
for (i in 1:length(mtab_targets)) {
	#this will find the index of the rows I want (based on mz values)
	r <- FindAllWithinError(oneList,maxErr,mtab_targets[i])

	if (length(r) > 0 ) {
		#show me the information for one compound...
		out <- pa@puritydf[r,]

		#in my example, out$seqNum is the scan number for the MS2 spectra
		#in my example, out$precursorScanNum or out$precursorNearest are the MS1 scans

		#so now, where the hell is the MS2 information from the purity results?
		#in the variable out I just created, the first column is 'pid' ...can use 
		#this to find the information about the grouped MS2 spectra in 
		#pa@grped_df
		#in my example, the first row in out has pid = 172256:
		#newdata <- pa@grped_df[which(pa@grped_df$pid==172256),]

		#can get all the MS2 from one group:
		#oneGroup <- pa@grped_df[which(pa@grped_df$grpid==753),]

		#then, once I have mapped the MS2 fragments to the MS1 data using 
		#frag4feature, I can (hopefully) map back to the original MS2 fragments
		#oneSet <- pa@grped_ms2$'753'

		tMean <- aggregate(out[,'iPurity'],list(out$filename),mean)
		idx <- match(tMean$Group.1,rownames(mean_iPurity))
		mean_iPurity[idx,i] <- tMean$x
		rm(tMean,idx)
		 
		#colnames(getMean) <- c('filename','mean_iPurity')
		#I think I finally have this...each row in oneGroup will have the MS2 scan
		#number given in 'precurMtchID' which is actually the MS2 scan. This row
		#corresponds to the section within oneSet. So, for row #1 in oneGroup I can
		#go to row #1 of oneSet to actually see the MS2 spectra.

		#plot up one set of iPurity scores for this mtab only
		#try ggplot
		#jpeg('test.jpg')
		#jpeg(paste0('iPurityPlot_',pos_mtabNames[i],'.jpg'))
		p10 <- ggplot(out,aes(x=out$filename,y=out$iPurity)) + geom_boxplot() 
		p10 <- p10 + ggtitle(paste0("iPurity, only for ", mtab_names[i])) 
		p10 <- p10 + theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
		print(p10)

		#dev.off()
	}
}


```


```{r iPurity_allStandards, eval = TRUE, fig.height = 8,fig.width = 10}

plot.new()
#par(mfrow=c(2,1)) 
par(mar = c(12, 4, 3, 3)) #bottom, left, top, right
#plot(df_score[,'taurocholic acid'],pch=16,col = useColors, cex=5)
boxplot(t(mean_iPurity),las=2,ylab = 'iPurity', 
	main = 'iPurity for all standards, dots are outliers')

```


```{r peaks_per_MS2, eval = TRUE, fig.height = 4,fig.width = 5}

#export the mz values: this does into puritydf (missed that yesterday)
oneList <- pa@puritydf[,'precursorMZ'] #data.frame syntax
maxErr <- 2
  
mean_ipkNm <- data.frame(matrix(ncol=length(mtab_targets),nrow=length(mzdatafiles)))
rownames(mean_ipkNm) <- basename(mzdatafiles)
colnames(mean_ipkNm) <- c(mtab_names)

#i <- 1
for (i in 1:length(mtab_targets)) {
	#this will find the index of the rows I want (based on mz values)
	r <- FindAllWithinError(oneList,maxErr,mtab_targets[i])

	if (length(r) > 0 ) {
		#show me the information for one compound...
		out <- pa@puritydf[r,]

		tMean <- aggregate(out[,'ipkNm'],list(out$filename),mean)
		idx <- match(tMean$Group.1,rownames(mean_ipkNm))
		mean_ipkNm[idx,i] <- tMean$x
		rm(tMean,idx)
		 
		p10 <- ggplot(out,aes(x=out$filename,y=out$ipkNm)) + geom_boxplot() 
		p10 <- p10 + ggtitle(paste0("ipkNm, only for ", mtab_names[i])) 
		p10 <- p10 + theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
		print(p10)

		#dev.off()
	}
}
```


```{r ipkNm_allStandards, eval = TRUE, fig.height = 8,fig.width = 10}

#plot.new()
#par(mfrow=c(2,1)) 
par(mar = c(12, 4, 3, 3)) #bottom, left, top, right
#plot(df_score[,'taurocholic acid'],pch=16,col = useColors, cex=5)
boxplot(t(mean_ipkNm),las=2,ylab = 'iPeak number (aka number of peaks in MS2 scan)', 
	main = 'mean iPeak Number for all standards, dots are outliers')

```


```{r nMS2_forEachStandard, eval = TRUE, fig.height = 4, fig.width = 5}

oneList <- pa@puritydf[,'precursorMZ'] #data.frame syntax
maxErr <- 2
  
number_MS2_perStd <- data.frame(matrix(ncol=length(mtab_targets),nrow=length(mzdatafiles)))
rownames(number_MS2_perStd) <- basename(mzdatafiles)
colnames(number_MS2_perStd) <- c(mtab_names)

#i <- 1
for (i in 1:length(mtab_targets)) {
	#this will find the index of the rows I want (based on mz values)
	r <- FindAllWithinError(oneList,maxErr,mtab_targets[i])

	if (length(r) > 0 ) {
		#show me the information for one compound...
		out <- pa@puritydf[r,]

		tData <- aggregate(out[,'pid'],list(out$filename),length)
		idx <- match(tData$Group.1,rownames(number_MS2_perStd))
		number_MS2_perStd[idx,i] <- tData$x
		rm(tData,idx)
		 
	}
}


```


```{r numberMS2_perStandard, eval = TRUE, fig.height = 8,fig.width = 10}

##leave the blanks out of this bc they have lots of MS2 scans for the standards

z <- intersect(rownames(number_MS2_perStd),paste0(putDataHere$File.Name,'.mzML'))
#want the index into rownames(df) 
idx1 <- match(z,rownames(number_MS2_perStd))
#then find the index into putDataHere
idx2 <- match(z,paste0(putDataHere$File.Name,'.mzML'))

number_MS2_perStd[idx1,'collision'] <- putDataHere$cType[idx2]
number_MS2_perStd[idx1,'energy'] <- putDataHere$cEnergy[idx2]
number_MS2_perStd[idx1,'detection'] <- putDataHere$dType[idx2]


#need reshape2 to use melt
number_MS2_perStd$id = rownames(number_MS2_perStd) #need this to keep the file names
toPlot2<- melt(number_MS2_perStd,value.name = 'nMS2_perStd',variable.name = 'standard')

#make the plot with ggplot
m1 <- ggplot(toPlot2,aes(x=toPlot2$id, y=toPlot2$nMS2_perStd,colour = toPlot2$standard)) +
    geom_point(aes(shape = toPlot2$detection), size=6)
m1 <- m1 + scale_color_brewer(palette = 'Paired')
m1 <- m1 + ggtitle('nMS2_perStd')
m1 <- m1 + theme_bw()
m1 <- m1 +  theme(plot.background = element_blank(),
   panel.grid.major = element_blank())
m1 <- m1 + theme(axis.text.x=element_text(angle = -45, hjust = 0))
print(m1)


```


##compMS2Miner results: scores from the in silico fragmentation

```{r compResults,eval = TRUE, fig.height = 8,fig.width = 10}

df_score <- data.frame(matrix(ncol=length(mtab_targets),nrow=length(mzdatafiles)))
rownames(df_score) <- basename(mzdatafiles)
colnames(df_score) <- c(mtab_names)

df_nHits <- df_score


for (im in 1:length(mzdatafiles)) {
	#get the oneComp_2 variable for each file...
	load(paste0(basename(mzdatafiles[im]),'_mined.RData'))
	
	#pull out the pieces I need for each search
	compMS2_object <- oneComp_2$oneComp
	peakTable <- oneComp_2$peakTable
	
	oneList <- peakTable$mzmed
	maxErr <- 4
	
	for (i in 1:length(mtab_names)) {  
		r <- FindAllWithinError(oneList,maxErr,mtab_targets[i])
		#make the string:
		##careful...will show multiple matches, which is good and bad
		#add steps to through one at a time 10/23/2017
		tFile <- paste0('CC_',r)
		if (length(tFile) > 1) {
		  #gather up the multiple scores...average them?
		  nH = length(tFile)
		  df_nHits[im,i] = nH
		  #go through one at time
		  temp_score <- data.frame(matrix(ncol=length(mtab_targets),nrow=nH))
		  colnames(temp_score) <- c(mtab_names)
		  
		  for (ih in 1:nH) {
      		#then use the string to search in compMS2_object for the results
      		#look for MetFrag hits in the compMS2Miner object
      		f <- paste0('compMS2_object@inSilico$MetFrag$',tFile[ih],'$DBname')
      		fs <- paste0('compMS2_object@inSilico$MetFrag$',tFile[ih])
      
      		names = eval(parse(text=f))
      		
      		if (is.null(names)){
      			#message(paste0(mtab_names[i],': no match')) 
      		} else {
      			fn <- match(mtab_names[i],tolower(names))
      			ot <- eval(parse(text = fs))
      			#then pull the score and ranked match from the 'ot' table
      			temp_score[ih,mtab_names[i]] = ot[fn,'Score']
      			
      		} #close searching for names
		    }
		  #try averaging the values...
		  d <- colMeans(temp_score,na.rm=TRUE)
		  df_score[im,mtab_names[i]] <- d[mtab_names[i]]
		  rm(d)
		  
		} else {
		  
  		#then use the string to search in compMS2_object for the results
  		#look for MetFrag hits in the compMS2Miner object
  		f <- paste0('compMS2_object@inSilico$MetFrag$',tFile,'$DBname')
  		fs <- paste0('compMS2_object@inSilico$MetFrag$',tFile)
  
  		names = eval(parse(text=f))
  		
  		if (is.null(names)){
  			#message(paste0(mtab_names[i],': no match')) 
  		} else {
  			fn <- match(mtab_names[i],tolower(names))
  			ot <- eval(parse(text = fs))
  			#then pull the score and ranked match from the 'ot' table
  			df_score[im,mtab_names[i]] = ot[fn,'Score']
  			df_nHits[im,mtab_names[i]] = 1
  		} #close searching for names
		}
		
	}

}

#add in the other variables before I do the plotting
z <- intersect(rownames(df_score),paste0(putDataHere$File.Name,'.mzML'))
#want the index into rownames(df) 
idx1 <- match(z,rownames(df_score))
#then find the index into putDataHere
idx2 <- match(z,paste0(putDataHere$File.Name,'.mzML'))

df_score_only <- df_score #use for averaging

df_score[idx1,'collision'] <- putDataHere$cType[idx2]
df_score[idx1,'energy'] <- putDataHere$cEnergy[idx2]
df_score[idx1,'detection'] <- putDataHere$dType[idx2]

toPlot_scores <- df_score #keep separate to get good plot in the rest of this chunk
df_score[idx1,'concStd'] <-putDataHere$stdConc_ng_per_ml[idx2]


#need reshape2 to use melt
toPlot_scores$id = rownames(toPlot_scores) #need this to keep the file names
toPlot<- melt(toPlot_scores,value.name = 'MetFragScore',variable.name = 'standard')


m <- ggplot(toPlot,aes(x=toPlot$id, y=toPlot$MetFragScore,colour = toPlot$standard)) +
    geom_point(aes(shape = toPlot$detection), size=6)


m <- m + scale_color_brewer(palette = 'Paired')
m <- m + ggtitle('MetFrag score, standards')
m <- m + theme_bw()
m <- m +  theme(plot.background = element_blank(),
   panel.grid.major = element_blank())
m <- m + theme(axis.text.x=element_text(angle = -45, hjust = 0))
print(m)


```

```{r finalSave, eval = TRUE}

save.image(NameOfFile)
```


